### 2016-10-13 paper/main.tex
added update

%\documentclass[journal]{IEEEtran}
\documentclass[journal,12pt,onecolumn,draftclsnofoot,]{IEEEtran}

\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{microtype}

\usepackage[cmex10]{amsmath}

\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}

% comment this out when you submit

\title{Duplicate Question Detection using Online Learning}
%\author{\IEEEauthorblockN{Author One\IEEEauthorrefmark{1},
%Author Two\IEEEauthorrefmark{2}, Author Three\IEEEauthorrefmark{3} and
%Author Four\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{Department of Whatever,
%Whichever University\\
%Wherever\\
%Email: \IEEEauthorrefmark{1}author.one@add.on.net,
%\IEEEauthorrefmark{2}author.two@add.on.net,
%\IEEEauthorrefmark{3}author.three@add.on.net,
%\IEEEauthorrefmark{4}author.four@add.on.net}}

\author{\IEEEauthorblockN{Chapman Siu}
\IEEEauthorblockA{\\College of Computing, 
Georgia Institute of Technology\\
Atlanta, GA 30332-0280, USA\\
Email: chapmansiu@gatech.edu}}

\maketitle

\begin{abstract}
%\boldmath
An integral part of OMSCS learning experience are online communities such as Piazza. Piazza is described as a ``Q\&A'' forum which is created for every class\cite{kincaid}. An integral part of many online communities is the concept of ``voting''' to determine popular and useful information. However, Piazza lacks this particular functionality which is on other Q\&A boards like Stackoverflow or Quora.

Being able to understand quality posts can be beneficial for students and instructors, as it will reduce the clutter in the community by allow students to ``see''' important posts in the follow up section of different threads and for instructors to identify potential experts and teaching assistant candidates for future classes. By building a corpus of high quality questions, this could even be extended to develop answers to user questions.

In this project, we aim to develop suitable approach for understanding quality of posts through introducing online machine learning techniques to assist with reducing clutter within young online communities.

\end{abstract}

\begin{IEEEkeywords}
duplicate question, Piazza
\end{IEEEkeywords}

\section{Introduction}

To understand and be able to explain quality posts, the type of modelling we are after is explanatory modelling, where we seek to provide causal explanations. One key consideration is that Shmueli\cite{modpred} describes how explanatory modelling does not always generalise well to predictive models. Shumueli\cite{modpred} also notes that the choice between explanatory and predictive modelling may affect the type of variables you use (variables have to make sense from a causal perspective), the type of preprocessing you may perform (data compression methods such as SVD may be inappropriate from a causal perspective), types of models you may build (ensemble of models would be inappropriate from a causal perspective).

Comparing with previous studies, we are aiming to demonstrate how these techniques can be used in an online learning context. Online learning focuses on modelling data which arrives in sequential order to update our best predictor for future data as opposed to training on the whole training data set at once. This is important part of this project as with young communities, often the whole training data set will not be available until the community is relatively mature. Pal \emph{et al.}\cite{pal} explains that early identification and engagement with these users  can improve the experience of these users and more importantly improve the overall quality of participation within a community. Within Pal \emph{et al.}\cite{pal} retrospective analysis, it had already shown that some of the potential experts had already left the community when the analysis concluded leading to missed opportunities within the community.

Furthermore to understand the type of questions which a community is asking can allow the noise to be diminished. For example on roughly 13 October 2016 the OMSCS admission results for Spring 2017 intake was annouced. Out of the newest 15 posts, 9 of them were new students questions or notes, with two threads having over 60 posts altogether. Similarly, reddit had 5 of 7 posts within the last 24 hours about new admissions, with 42 out of 44 comments in the respective posts in the threads on admissions.

Having adaptive duplication detection not only assists in special events but also in young communities, where there is a lack of training data or examples to create a sensible detection for duplicated questions. On forums like Piazza this would assist in reducing the amount of clutter which exists, similar to how stackoverflow closes duplicate questions for redundant questions to decrease the maintenance and people's resources on answering the same question.

The following usage scenarios demonstrate the benefits of such a tool.

\emph{Scenario} 1 - \emph{Without tool.} Steve recently gained admission to OMSCS program. However he is confused by the registration process and is unsure which course he should pick. He posts to the Google Plus community but received no response due to the flood of registration related posts.

\emph{Scenario} 2 - \emph{With tool.} Steve recently gained admission to OMSCS program. However he is confused by the registration process and is unsure which course he should pick. He posts to the Google+ community. By using our automated tool, we can readily detect similar questions that have been asked before and direct Steve to a useful resource.

We evaluate our approach on synthetically generated posts based on real questions asked on Google+ and Reddit, combined with unofficial FAQs generated by the community. We will also evaluate the performance by simulating a newly created Piazza forum to see if the tool has a similar efficacy.

The main contributions of this paper are as follows.

\begin{itemize}
    \item We propose the problem of duplicate question detection in online communities. We procose a novel approach which considers and integrates multiple factors to detect duplicate questions.
    \item We evalute different types of communities based on real and synthetic questions.
\end{itemize}

The remainder of the paper is organised as follows. We elaborate on the motivation of our work and introduce online learning variants of Latent Dirichlet Allocation (LDA) and Word2Vec, and describe the other components of the overall framework. Next we will discuss some issues about the performance, efficiency and threats to validity. We will review related work, conclude the paper and mention future work.

\section{Preliminaries}

%online Latent Dirichlet Allocation (online-LDA), which is used to extract topics from a young community in an iterative way and gain topic distribution from natural language description of questions.

%Online word2vec algorithm - word to vec is...

%\subsection{Motivation}

\subsection{Online Latent Dirichlet Allocation}

Online Latent Dirichlet Allocation is an online variant of Latent Dirichlet Allocation (LDA). LDA is a well-known topic modelling technique proposed by Blei \emph{et al.}\cite{lda}. LDA is a generative probabilistic model of a textual corpus (i.e., a set of textual documents), which takes a training textual corpus as input, and a number of parameters including the number of topics $(K)$ considered. In the training phase, for each document $s$, LDA will compute its topic distribution $\theta_s$, which is a vector with $K$ elements, and each element corresponds to a topic. The value of each element in $\theta_s$ is a real number from $0$ to $1$, which represents the proportion of the words in $s$ that belong to the corresponding topic in $s$. After training, LDA can be used to predict the topic distribution $\theta_m$ of a new document $m$. In our case, a document is the description of a question, and the topic is a higher level concept corresponding to a distribution of words. For example, we may have the topic "admissions", which is a distribution of words such as ``citizenship'', ``GRE'', ``TOEFL'', ``transcripts''.

LDA can be extended in an online learning problem by reframing LDA using approximate inference techniques (variational inference) which then becomes an optimization problem and allows LDA to be trained via known approaches such as stochastic gradient descent\cite{onlinelda}. % explain this more...

% intuitively it means we find a problem which is easier to solve and show that it "converges in distribution" or close enough - according to KL divergence
% and based on this we can use gradient descent and by extension mini-batch+stochastic extensions. (left without proof?)

\subsection{Online Word2Vec}

Word2Vec is all about computing distributed vector representations of words. In this project we will be using the skip-gram variant.

% from spark doco

The training objective of skip-gram is to learn word vector representations that are good at predicting its context in the same sentence. Mathematically, given a sequence of training words $w_1,w_2,...,w_T$, the objective of the skip-gram model is to maximize the average log-likelihood

$$ \frac{1}{T}\sum_{t=1}^T\sum_{j=-k}^{j=k} \log \Pr(w_{t+j} | w_t) $$

where $k$ is the size of the training window.

In the skip-gram model, every word $w$ is associated with two vectors $u_w$ and $v_w$ which are vector representations of $w$ as word and context respectively. The probability of correctly predicting word $w_i$ given word $w_j$ is determined by the softmax model, which is

$$ \Pr (w_i | w_j) = \frac{\exp(u_{w_i}^T v_{w_j})}{\sum_{l=1}^V \exp(u_l^T v_{w_j}} $$

where $V$ is the vocabulary size.

The skip-gram model with softmax is expensive because the cost of computing $\log (\Pr(w_i | w_j))$ is proportional to $V$, which can be easily in order of millions.% as stated in gensim negative sampling is used.- what does richard say about how negative sampling works?

% (see tensorflow, spark page) - can copy spark page for this.
% copy spark page
% http://spark.apache.org/docs/latest/mllib-feature-extraction.html#word2vec

Online variant of Word2Vec solves the online learning problem, by updating vocabulary whenever new documents are ingested. The new words are then initialized with random weights, whilst existing words retain their weights as normal. Then training will be again be initialised in an iterative fashion. % cite this?

\subsection{Cosine Similarity}

%(Write Why we don't need online learning here)

Cosine similarity is used to compute the similiarty between pairs of sets of words based on common words that they share. After preprocessing, the words are transformed into two bags (i.e. multisets) of words. For two sets of words $m$ and $n$, we represent the two bags of words that extracted as $Bag_m$ and $Bag_n$ respectively. Next we merge $Bag_m$ and $Bag_n$ and eliminate duplicate words to obtain the union set $Bag_u$, which contains $v$ words.  Following vector space modelling, we represent the two sentences as two vectors: $Vec_m = (wt_{m,1}, wt_{m, 2}, ..., wt_{m, v})$ and $Vec_n = (wt_{n,1}, wt_{n, 2}, ..., wt_{n, v})$. The weight $wt_{q, i}$  denotes the relative term frequency of the $i$-th word in sentence $q$'s title, which is computed as follows:

$$ wt_{q, i} = \frac{n_{q, i}}{\sum_v n_{q, v}} $$

where $n_{q, i}$ denotes the number of times the $i$-th word of $Bag_u$ appears in the sentence $q$, $\sum_v n_{q,v}$ denotes the total number of occurences of all words in the title of question $q$, where $v$ is the index of the word in $Bag_u$. We measure the similarity between two questions' titles by computing the cosine similarity of their vector representations $Vec_m$ and $Vec_n$ as follows:

$$ CosineSim(Vec_m, Vec_n) = \frac{\langle Vec_m,  Vec_n \rangle}{|Vec_m| | Vec_n| }$$

The numerator $\langle Vec_m , Vec_n \rangle$ which is the dot product of the two vectors

$$ \langle Vec_m , Vec_n \rangle  = wt_{m, 1} \times wt_{n, 1} + ... + wt_{m, v} wt_{n, v} $$

The terms $|Vec_m|$ and $|Vec_n|$ in the denomiator denote the sizes of the two vectors respectively, where the size of $Vec_m$ is computed as

$$ | Vec_m  | = \sqrt{wt_{m,1}^2 + wt_{m,2}^2 + ... + wt_{m,v}^2} $$

Cosine similarity measures do not require the whole corpus in order to compute similarity; rather it depends only on the pairwise sets of words. As such there is no online variant to cosine similarity as one does not need to build or store a corpus in its formulation.

\section{Proposed Approach}

\subsection{Overall Framework}

Insert some image of the framework and description of how it would work overtime.

\subsection{Feature Construction}

Probably revolves around the more qualitative side, as described here: 
\url{http://researcher.ibm.com/researcher/files/us-apal/umap11_earlyexperts.pdf}

\subsection{Estimation of Weights}

Logistic regression would probably be good enough...

\section{Experiments and Results}

\subsection{Experimental Setup}

Based on synthetic, Piazza, Google plus.

\subsection{Evaluation Metrics}

Which one to use and why?

\subsection{Research Questions and Findings}

How effective? How does it compare to a batch learning approach.

\section{Discussion}

\subsection{Threats to Validity}

Threats include the ability of this approach to generalise to all young communities. As young communities is quite broad, the focus in this paper has been quite targeted, looking at computer science forums, and learning centers. The duplicate questions were also quite targeted and manually pruned. In an more unbalanced set, this ....?

Threats to construct validity - suitability of evaluation metrics

\section{Related Work}

Besides what has been in the proposal, links to online learning, such as vowpal wabbit:

\section{Conclusion}

Some conclusion

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Some text for the appendix.

% use section* for acknowledgement

\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\bibitem{kincaid}
Kincaid, J. (2010, February 5). ``Piazzza gives classmates an online forum to trade their knowledge'''. Retrieved September 18, 2016, from \url{https://techcrunch.com/2010/02/05/piazzza-college-questions-answers/}

\bibitem{pal}
Pal, A., Farzan, R., Konstan, J. A., and Kraut, R. E. ``Early detection of potential experts in question answering communities.'' in \emph{User Modeling, Adaption and Personalization}. Springer, 2011, 231-242. Retrieved on 12 September 2016 from \url{http://researcher.ibm.com/researcher/files/us-apal/umap11_earlyexperts.pdf}

\bibitem{modpred}
Shmueli, G. (2010). ``To explain or to predict?'' \emph{Statistical Science} 25(3), 289–310. doi:10.1214/10-sts330. Retrieved on 30 September from \url{http://www.stat.berkeley.edu/~aldous/157/Papers/shmueli.pdf}

\bibitem{introIR}
Manning C.D., Raghavan P. and Schütze H. (2008), ``Introduction to Information Retrieval'', Cambridge University Press. 2008. Retreived on 30 September from \url{http://www-nlp.stanford.edu/IR-book/}

\bibitem{bog}
Bogatyy I. (2016) ``Predicting answer types for question-answering''. Retrieved on 30 September from \url{https://cs224d.stanford.edu/reports/Bogatyy.pdf}

\bibitem{dupred}
Zhang, Y., Lo, D., Xia, X., and Sun, J.-L. (2015). ``Multi-factor duplicate question detection in stack overflow'. \emph{Journal of Computer Science and Technology}, 30(5), 981–997. Retrieved on 17 September 2016 from \url{https://soarsmu.github.io/papers/jcst-duplicateqns.pdf}.  doi:10.1007/s11390-015-1576-4

\bibitem{lda}
Blei, D. M., Ng, A. Y. and Jordan, M. I. (2003). ``Latent dirichlet allocation''. J. Mach. Learn. Res., 3, 993--1022. doi: 10.1162/jmlr.2003.3.4-5.993. Retreived on 12 October from \url{www.jmlr.org/papers/volume3/blei03a/blei03a.pdf}

\bibitem{onlinelda}
Hoffman, M. D., Blei, D. M. and Bach F. (2010). ``Online Learning for Latent Dirichlet Allocation'' Retreived on 12 October from \url{http://www.cs.princeton.edu/~blei/papers/HoffmanBleiBach2010b.pdf}

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{Chapman Siu}
is Masters candidate from Georgia Institute of Technology. Chapman is also a lion.
\end{IEEEbiography}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% comment this out when you submit

% that's all folks
\end{document}